Bandit Multiclass Linear Classification: \\ Efficient Algorithms for the Separable Case

### 2019-05-03 ICML-camera-ready/appendix-experiment.tex
entomological changes; reviews; some formatting issues fixed

We plot in
Figures~\ref{fig:banditron-points},~\ref{fig:linearova-points},~\ref{fig:rationalova-points}
the final decision boundaries learned by each algorithm in the two datasets (Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}), i.e. its decision boundaries
output at the end of round $T = 5 \times 10^6$.

>>>>

Figures~\ref{fig:banditron-points},~\ref{fig:linearova-points}, and~\ref{fig:rationalova-points} show the final decision boundaries learned by each algorithm on the two datasets (Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset} and~\ref{figure:number-of-mistakes-weakly-separable-dataset}), 
after $T = 5 \times 10^6$ rounds. We used the version of Banditron
with exploration rate of 0.02, which explores the most.

### 2019-05-03 ICML-camera-ready/empirical-comparison.tex
entomological changes; reviews; some formatting issues fixed

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with rational
kernel~\eqref{equation:rational-kernel} and used implementation of
\textsc{Banditron} algorithm by \citet{Orabona09}.\footnote{The source code
is included in the supplementary material.}
We evaluated these algorithms on the two datasets. \textsc{Banditron} has an
exploration rate parameter, for which we tried values $0.02, 0.01, 0.005, 0.002,
0.001, 0.0005$. Since all three algorithms are randomized, we run each algorithm
$20$ times. The average cumulative number of mistakes up to round $t$ as a
function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

@@

\begin{minipage}{.45\textwidth}
\begin{figure}[H]
\centering
\includegraphics[width=0.98\textwidth]{figures/strong3}
\caption{Average cumulative number of mistakes of various algorithms versus the
number of rounds for the strongly linearly separable dataset of
\autoref{figure:strongly-separable-dataset}.}
\label{figure:number-of-mistakes-strongly-separable-dataset}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\begin{figure}[H]
\centering
\includegraphics[width=0.98\textwidth]{figures/weak3}
\caption{Average cumulative number of mistakes of various algorithms versus the
number of rounds for the weakly linearly separable dataset of
\autoref{figure:weakly-separable-dataset}.}
\label{figure:number-of-mistakes-weakly-separable-dataset}
\end{figure}
\end{minipage}

>>>>

We implemented
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples},
Algorithm~\ref{algorithm:kernelized} with rational
kernel~\eqref{equation:rational-kernel} and used implementation of
\textsc{Banditron} algorithm by \citet{Orabona09}.%\footnote{The source code
%is included in the supplementary material.}
We evaluated these algorithms on the two datasets. \textsc{Banditron} has an
exploration rate parameter, for which we tried values $0.02, 0.01, 0.005, 0.002,
0.001, 0.0005$. Since all three algorithms are randomized, we run each algorithm
$20$ times. The average cumulative number of mistakes up to round $t$ as a
function of $t$ are shown in
Figures~\ref{figure:number-of-mistakes-strongly-separable-dataset}
and~\ref{figure:number-of-mistakes-weakly-separable-dataset}.

@@

\begin{minipage}{.45\textwidth}
\begin{figure}[H]
\centering
\includegraphics[width=0.98\textwidth]{figures/strong3}
\caption{The average cumulative number of mistakes versus the
number of rounds on the strongly linearly separable dataset in
\autoref{figure:strongly-separable-dataset}.}
\label{figure:number-of-mistakes-strongly-separable-dataset}
\end{figure}
\end{minipage}
\hfill
\begin{minipage}{.45\textwidth}
\begin{figure}[H]
\centering
\includegraphics[width=0.98\textwidth]{figures/weak3}
\caption{The average cumulative number of mistakes versus the
number of rounds on the weakly linearly separable dataset in
\autoref{figure:weakly-separable-dataset}.}
\label{figure:number-of-mistakes-weakly-separable-dataset}
\end{figure}
\end{minipage}

### 2019-05-03 ICML-camera-ready/from-weak-separability-to-strong-separability.tex
entomological changes; reviews; some formatting issues fixed

We use a positive definite kernel function $k(\cdot, \cdot)$, namely
\emph{rational kernel}~\citep{Shalev-Shwartz-Shamir-Sridharan-2011}, whose
corresponding feature map $\phi(\cdot)$ transforms any sequence of \emph{weakly}
linearly separable examples to \emph{strongly} linearly separable sequence of
examples.
Specifically, $\phi$ has the property that if a set of labeled
examples in $\B(\zero,1)$ is weakly linearly separable with a margin $\gamma$, then
after applying $\phi$ the examples become strongly linearly separable with a
margin $\gamma'$ and their squared norms are bounded by $2$.
\footnote{Other kernels, such as the polynomial kernel
$k(x,x') = (1+\ip{x}{x'})^d$, or the multinomial kernel~\citep{Goel-Klivans-2017} $k(x,x') = \sum_{i=0}^d(\ip{x}{x'})^i$,
will have similar properties for large enough $d$.}
The parameter $\gamma'$ is
a function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

Consider the classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty
~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner
product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. If we index the
coordinates of $\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of
non-negative integers, the feature map that corresponds to $k$ is $\phi: \B(\zero,1)
\to \ell_2$,
\begin{align}
& \left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\label{equation:phi}
\end{align}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots,
\alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2!
\dots \alpha_d!}$ is the multinomial coefficient. It can be easily checked that
$$
k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}.
$$
The last equality together with the formula for $k$ implies that $k(x,x) <
+\infty$ for any $x$ in $\B(\zero,1)$ and thus in particular implies that $\phi(x)$
indeed lies in $\ell_2$.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ in $\B(\zero,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
%\begingroup
%\allowdisplaybreaks
%\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
%\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
%& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
%\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(4K-5) 2^{K-1}} \\
%& \ \cdot \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
%\end{align*}
%\endgroup
\begingroup
\allowdisplaybreaks
\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 = & \frac{
  \left[
    376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil
  \right]^{
    \frac{-\lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{{2}/{\gamma}} \right \rceil}{2}
  }
}{2\sqrt{K}},
\gamma_2 = & \frac{
   \left(2^{s+1} r(K-1) (4s+2) \right)^{-(s+1/2)r(K-1)}
}{4\sqrt{K}(4K-5) 2^{K-1}}
\; ,
\end{align*}
\endgroup

@@

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, \qquad p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t}, \qquad
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with a positive margin.

>>>>

We use a positive definite kernel function $k(\cdot, \cdot)$, namely
a \emph{rational kernel}~\citep{Shalev-Shwartz-Shamir-Sridharan-2011} whose
corresponding feature map $\phi(\cdot)$ transforms any sequence of \emph{weakly}
linearly separable examples to a \emph{strongly} linearly separable sequence of
examples.
Specifically, $\phi$ has the property that if a set of labeled
examples in $\B(\zero,1)$ is weakly linearly separable with a margin $\gamma$, then
after applying $\phi$ the examples become strongly linearly separable with a
margin $\gamma'$ and their squared norms are bounded by $2$.
\footnote{Other kernels, such as the polynomial kernel
$k(x,x') = (1+\ip{x}{x'})^d$, or the multinomial kernel~\citep{Goel-Klivans-2017} $k(x,x') = \sum_{i=0}^d(\ip{x}{x'})^i$,
will have similar properties for large enough $d$.}
The parameter $\gamma'$ is
a function of the old margin $\gamma$ and the number of classes $K$, and is
specified in \autoref{theorem:margin-transformation} below.

@@

Consider the classical real separable Hilbert space $\ell_2 = \{ x \in \R^\infty
~:~ \sum_{i=1}^\infty x_i^2 < + \infty \}$ equipped with the standard inner
product $\ip{x}{x'}_{\ell_2} = \sum_{i=1}^\infty x_i x'_i$. If we index the
coordinates of $\ell_2$ by $d$-tuples $(\alpha_1, \alpha_2, \dots, \alpha_d)$ of
non-negative integers, the feature map that corresponds to $k$ is $\phi: \B(\zero,1)
\to \ell_2$,
\begin{align}
\begin{split}
&\left(\phi(x_1, x_2, \dots, x_d)\right)_{(\alpha_1, \alpha_2, \dots, \alpha_d)} = \\
& x_1^{\alpha_1} x_2^{\alpha_2} \dots x_d^{\alpha_d} \cdot \sqrt{2^{-(\alpha_1 + \alpha_2 + \dots + \alpha_d)} \binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots, \alpha_d}}
\end{split}
\label{equation:phi}
\end{align}
where $\binom{\alpha_1 + \alpha_2 + \dots + \alpha_d}{\alpha_1, \alpha_2, \dots,
\alpha_d} = \frac{(\alpha_1 + \alpha_2 + \dots + \alpha_d)!}{\alpha_1! \alpha_2!
\dots \alpha_d!}$ is the multinomial coefficient. It can be easily checked that
$$
k(x,x') = \ip{\phi(x)}{\phi(x')}_{\ell_2}.
$$
The last equality together with the formula for $k$ implies that $k(x,x) <
+\infty$ for any $x$ in $\B(\zero,1)$ and thus in particular implies that $\phi(x)$
indeed lies in $\ell_2$.

@@

\begin{theorem}[Margin transformation]
\label{theorem:margin-transformation}
Let $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_T, y_T)$ $\in \B(\zero,1) \times
\{1,2,\dots,K\}$ be a sequence of labeled examples that is weakly linearly
separable with margin $\gamma > 0$. Let $\phi$ be as defined in
equation~\eqref{equation:phi} and let
%\begingroup
%\allowdisplaybreaks
%\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
%\gamma_1 & = \frac{1}{2\sqrt{K}}  \\
%& \ \cdot \left(376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil \right)^{-\frac{1}{2} \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil}, \\
%\gamma_2 & = \frac{2^{s(s+1)r(K-1)} }{4\sqrt{K}(4K-5) 2^{K-1}} \\
%& \ \cdot \left(2^{s+1} r(K-1) (4s+2)^2 \right)^{-(s+1/2)r(K-1)} \; .
%\end{align*}
%\endgroup
\begingroup
\allowdisplaybreaks
\begin{align*}
%r & = 2 \left\lceil \frac{1}{4} \log_2(4K-3) \right\rceil + 1, \quad \quad s = \left \lceil \log_2(2/\gamma) \right \rceil, \\
\gamma_1 = & \frac{
  \left[
    376 \lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{\frac{2}{\gamma}} \right \rceil
  \right]^{
    \frac{-\lceil \log_2(2K-2) \rceil \cdot \left \lceil \sqrt{{2}/{\gamma}} \right \rceil}{2}
  }
}{2\sqrt{K}},\\
\gamma_2 = & \frac{
   \left(2^{s+1} r(K-1) (4s+2) \right)^{-(s+1/2)r(K-1)}
}{4\sqrt{K}(4K-5) 2^{K-1}}
\; ,
\end{align*}
\endgroup

@@

\paragraph{Overview.} The idea behind the construction and analysis of the
mapping $\phi$ is polynomial approximation. Specifically, we construct $K$
multivariate polynomials $p_1, p_2, \dots,p_K$ such that
\begin{align}
\label{equation:poly-pos}
& \forall t \in \cbr{1,2,\dots,T}, \qquad p_{y_t}(x_t) \ge \frac{\gamma'}{2} \; ,
\\
\label{equation:poly-neg}
& \begin{gathered}
\forall t \in \cbr{1,2,\dots,T} \ \forall i \in \cbr{1,2,\ldots,K} \setminus \cbr{y_t}, \qquad
p_i(x_t) \le - \frac{\gamma'}{2} \; .
\end{gathered}
\end{align}
We then show (\autoref{lemma:norm-bound}) that each polynomial $p_i$ can be
expressed as $\ip{c_i}{\phi(x)}_{\ell_2}$ for some $c_i \in \ell_2$. This
immediately implies that the examples $(\phi(x_1),y_1), \ldots,
(\phi(x_T),y_T)$ are strongly linearly separable with a positive margin.

### 2019-05-03 ICML-camera-ready/introduction.tex
entomological changes; reviews; some formatting issues fixed

The bandit feedback setting, however, is much more challenging. For the case
when the examples are strongly linearly separable, to the best of our knowledge,
it is not known how to design an efficient algorithm with a finite mistake bound
before our work.~\footnote{Although~\cite{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their
Theorem~2 is incorrect: first, their Lemma~1 (with $C = +\infty$) along with their Theorem~1
implies a mistake upper bound of
$(\frac{R}{\gamma})^2$, which contradicts the lower
bound in our Theorem~\ref{theorem:strongly-separable-examples-mistake-lower-bound};
second, their Lemma~1 cannot be directly applied to the bandit feedback setting.}
We design a simple and
efficient algorithm
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
that makes at most $O(K (R/\gamma)^2)$ mistakes in expectation. Its memory
complexity and per-round time complexity are both $O(dK)$. The algorithm can be
viewed as running $K$ copies of the \textsc{Binary Perceptron} algorithm, one
copy for each class. We prove that any (possibly randomized) algorithm must make
$\Omega(K (R/\gamma)^2)$ mistakes in the worst case. The extra $O(K)$
multiplicative factor in the mistake bound, as compared to the full-information
setting, is the price we pay for the bandit feedback, or more precisely, the
lack of full-information feedback.

>>>>

The bandit feedback setting, however, is much more challenging. For the 
strongly linearly separable case, we are not aware of any 
prior efficient algorithm with a finite mistake bound.
~\footnote{Although~\citet{Chen-Chen-Zhang-Chen-Zhang-2009}
claimed that their Conservative OVA algorithm with PA-I update has a finite
mistake bound under the strong linear separability condition, their
Theorem~2 is incorrect: first, their Lemma~1 (with $C = +\infty$) along with their Theorem~1
implies a mistake upper bound of
$(\frac{R}{\gamma})^2$, which contradicts the lower
bound in our Theorem~\ref{theorem:strongly-separable-examples-mistake-lower-bound};
second, their Lemma~1 cannot be directly applied to the bandit feedback setting.}
We design a simple and
efficient algorithm
(Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples})
that makes at most $O(K (R/\gamma)^2)$ mistakes in expectation. Its memory
complexity and per-round time complexity are both $O(dK)$. The algorithm can be
viewed as running $K$ copies of the \textsc{Binary Perceptron} algorithm, one
copy for each class. We prove that any (possibly randomized) algorithm must make
$\Omega(K (R/\gamma)^2)$ mistakes in the worst case. The extra $O(K)$
multiplicative factor in the mistake bound, as compared to the full-information
setting, is the price we pay for the bandit feedback, or more precisely, the
lack of full-information feedback.

### 2019-05-03 ICML-camera-ready/mistake-lower-bound-for-ignorant-algorithms.tex
entomological changes; reviews; some formatting issues fixed

To explain the definition, consider an ignorant algorithm $\calA$. Suppose that
on a sequence of examples $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_{t-1}, y_{t-1})$
generated by some adversary the algorthm $\calA$ makes correct predictions in
rounds $a_1, a_2, \dots, a_n$ where $1 \le a_1 < a_2 < \dots < a_n < t$ and
errors on rounds $\{1,2,\dots,t-1\} \setminus \{a_1, a_2, \dots, a_n\}$. Suppose
that on another sequence of examples $(x_1', y_1'), (x_2', y_2'), \dots,
(x_{s-1}', y_{s-1}')$ generated by another adversary the algorithm $\calA$ makes
correct predictions in rounds $b_1, b_2, \dots, b_n$ where $1 \le b_1 < b_2 <
\dots < b_n < s$ and errors on rounds $\{1,2,\dots,s-1\} \setminus \{b_1, b_2,
\dots, b_n\}$. Futhermore, suppose
\begin{align*}
(x_{a_1}, y_{a_1}) &= (x'_{b_1}, y'_{b_1}) \; , \\
(x_{a_2}, y_{a_2}) &= (x'_{b_2}, y'_{b_2}) \; , \\
\vdots \\
(x_{a_n}, y_{a_n}) &= (x'_{b_2}, y'_{b_n}) \; .
\end{align*}
Then, as $\calA$ is ignorant,
\[
\Pr[y_t = y ~|~ (x_1, y_1), (x_2, y_2) \dots, (x_{t-1}, y_{t-1}), x_t = x] =
\Pr[y_t' = y ~|~ (x_1', y_1'), (x_2', y_2') \dots, (x_{t-1}', y_{t-1}'), x_t' = x] \; .
\]
Note that the sequences $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_{t-1},
y_{t-1})$ and $(x_1', y_1')$, $(x_2', y_2')$, $\dots$, $(x_{s-1}', y_{s-1}')$
might have different lengths and and $\calA$ might error in different sets of
rounds. As a special case, if an ignorant algorithm makes a mistake in round $t$
then $p_{t+1}=p_t$.

@@

In rounds $1,2,\dots,\tau-1$, the algorithm predicts the incorrect class $2$
with probability at least $q_0$. Thus,
\begin{equation}
\Exp\left[\sum_{t=1}^{\tau - 1} \indicator{\widehat y_t\neq y_t}\right] = q_0 \Exp[(\tau - 1)] \; .
\label{eqn:err-phase1}
\end{equation}
In rounds $\tau, \tau+1, \dots, T$, all the examples are the same and are equal
to $(v_\tau, 2)$. Let $s$ be the first time step $t$ such that $t \ge \tau$
and the algorithm makes a correct prediction. If the algorithm makes mistakes
in all rounds $\tau, \tau+1, \dots, T$, we define $s = T+1$.
By definition the algorithm makes mistakes in rounds $\tau, \tau+1, \dots, s-1$.
Therefore,
\begin{equation}
\Exp\left[\sum_{t=\tau}^T \indicator{\widehat y_t\neq y_t}\right] \ge \Exp[s-\tau].
\label{eqn:err-phase2}
\end{equation}
%Since the algorithm is ignorant, conditioned on $\tau$, $s-\tau+1$ has the same distribution as
%$\min(X, T-\tau+2)$, where $X$ is drawn from a geometric distribution with parameter
%$p_\tau(2|v_\tau)$.
Since the algorithm is ignorant, conditioned on $\tau$ and $q \triangleq p_\tau(2|v_\tau)$, $s-\tau$ follows a truncated geometric distribution with parameter $q$
(i.e., $s-\tau$ is $0$ with probability $q$, $1$ with probability $(1-q)q$, $2$ with probability $(1-q)^2q, \ldots$). Its conditional expectation can be calculated as follows:
\begin{align*}
\Exp[s-\tau~|~\tau, q]
&= \sum_{i=1}^{T+1-\tau} i \times \Pr[s-\tau = i |~\tau, q ] \\
&= \sum_{j=1}^{T+1-\tau}\Pr[s-\tau\geq j | ~\tau, q] = \sum_{j=1}^{T+1-\tau}(1-q)^j \geq  \sum_{j=1}^{T+1-\tau}(1-q_0)^j = \frac{1-q_0}{q_0}\left( 1-(1-q_0)^{T-\tau+1} \right).
\end{align*}

>>>>

To explain the definition, consider an ignorant algorithm $\calA$. Suppose that
on a sequence of examples $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_{t-1}, y_{t-1})$
generated by some adversary the algorthm $\calA$ makes correct predictions in
rounds $a_1, a_2, \dots, a_n$ where $1 \le a_1 < a_2 < \dots < a_n < t$ and
errors on rounds $\{1,2,\dots,t-1\} \setminus \{a_1, a_2, \dots, a_n\}$. Suppose
that on another sequence of examples $(x_1', y_1'), (x_2', y_2'), \dots,
(x_{s-1}', y_{s-1}')$ generated by another adversary the algorithm $\calA$ makes
correct predictions in rounds $b_1, b_2, \dots, b_n$ where $1 \le b_1 < b_2 <
\dots < b_n < s$ and errors on rounds $\{1,2,\dots,s-1\} \setminus \{b_1, b_2,
\dots, b_n\}$. Futhermore, suppose
\begin{align*}
(x_{a_1}, y_{a_1}) &= (x'_{b_1}, y'_{b_1}) \; , \\
(x_{a_2}, y_{a_2}) &= (x'_{b_2}, y'_{b_2}) \; , \\
\vdots \\
(x_{a_n}, y_{a_n}) &= (x'_{b_2}, y'_{b_n}) \; .
\end{align*}
Then, as $\calA$ is ignorant,
\begin{align*}
\Pr[y_t = y ~|~ (x_1, y_1), (x_2, y_2) \dots, (x_{t-1}, y_{t-1}), x_t = x] = \\
\Pr[y_t' = y ~|~ (x_1', y_1'), (x_2', y_2') \dots, (x_{t-1}', y_{t-1}'), x_t' = x].
\end{align*}
Note that the sequences $(x_1, y_1)$, $(x_2, y_2)$, $\dots$, $(x_{t-1},
y_{t-1})$ and $(x_1', y_1')$, $(x_2', y_2')$, $\dots$, $(x_{s-1}', y_{s-1}')$
might have different lengths and and $\calA$ might error in different sets of
rounds. As a special case, if an ignorant algorithm makes a mistake in round $t$
then $p_{t+1}=p_t$.

@@

In rounds $1,2,\dots,\tau-1$, the algorithm predicts the incorrect class $2$
with probability at least $q_0$. Thus,
\begin{equation}
\Exp\left[\sum_{t=1}^{\tau - 1} \indicator{\widehat y_t\neq y_t}\right] = q_0 \Exp[(\tau - 1)] \; .
\label{eqn:err-phase1}
\end{equation}
In rounds $\tau, \tau+1, \dots, T$, all the examples are the same and are equal
to $(v_\tau, 2)$. Let $s$ be the first time step $t$ such that $t \ge \tau$
and the algorithm makes a correct prediction. If the algorithm makes mistakes
in all rounds $\tau, \tau+1, \dots, T$, we define $s = T+1$.
By definition the algorithm makes mistakes in rounds $\tau, \tau+1, \dots, s-1$.
Therefore,
\begin{equation}
\Exp\left[\sum_{t=\tau}^T \indicator{\widehat y_t\neq y_t}\right] \ge \Exp[s-\tau].
\label{eqn:err-phase2}
\end{equation}
%Since the algorithm is ignorant, conditioned on $\tau$, $s-\tau+1$ has the same distribution as
%$\min(X, T-\tau+2)$, where $X$ is drawn from a geometric distribution with parameter
%$p_\tau(2|v_\tau)$.
Since the algorithm is ignorant, conditioned on $\tau$ and $q \triangleq p_\tau(2|v_\tau)$, $s-\tau$ follows a truncated geometric distribution with parameter $q$
(i.e., $s-\tau$ is $0$ with probability $q$, $1$ with probability $(1-q)q$, $2$ with probability $(1-q)^2q, \ldots$). Its conditional expectation can be calculated as follows:
\begin{align*}
\Exp[s-\tau~|~\tau, q]
&= \sum_{i=1}^{T+1-\tau} i \times \Pr[s-\tau = i |~\tau, q ] \\
&= \sum_{j=1}^{T+1-\tau}\Pr[s-\tau\geq j | ~\tau, q] \\
& = \sum_{j=1}^{T+1-\tau}(1-q)^j \geq  \sum_{j=1}^{T+1-\tau}(1-q_0)^j \\
& = \frac{1-q_0}{q_0}\left( 1-(1-q_0)^{T-\tau+1} \right).
\end{align*}

### 2019-05-03 ICML-camera-ready/notions-of-linear-separability.tex
fix some formatting issues
entomological changes; reviews; some formatting issues fixed

\section{Notions of linear separability}
\label{section:notions-of-linear-separability}

We define two notions of linear separability for multiclass classification. The
first notion is the standard notion of linear separability used in the proof of
the mistake bound for the \textsc{Multiclass Perceptron} algorithm \citep[see e.g.][]{Crammer-Singer-2003}. The second
notion is stronger, i.e. more restrictive. %However, it is more suitable for the
%bandit setting, since it allows for a simple and efficient algorithm; see
%Section~\ref{section:algorithm-for-strongly-linearly-separable-data}.

\begin{definition}
[Linear separability]
\label{definition:linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number.
We say that labeled examples $(x_1, y_1),
(x_2, y_2), \dots, (x_T, y_T) \in V \times \{1,2,\dots,K\}$ are
%said to be
\begin{enumerate}
\item
\emph{weakly linearly separable with a margin $\gamma$} if there exist vectors
$w_1, w_2, \dots, w_K \in V$ such that
\begin{align}
\label{equation:weak-linear-separability-1}
& \sum_{i=1}^K \norm{w_i}^2 \le 1 \; , \\
\label{equation:weak-linear-separability-2}
& \begin{gathered}
\forall t \in \{1,2,\dots,T\} \quad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\} \qquad
\ip{x_t}{w_{y_t}} \ge \ip{x_t}{w_i} + \gamma \; ,
\end{gathered}
\end{align}
\item
\emph{strongly linearly separable with a margin $\gamma$} if there exist vectors
$w_1, w_2, \dots, w_K \in V$ such that
\begin{align}
\label{equation:strong-linear-separability-1}
& \sum_{i=1}^K \norm{w_i}^2 \le 1 \; , \\
\label{equation:strong-linear-separability-2}
& \forall t \in \{1,2,\dots,T\} \qquad \ip{x_t}{w_{y_t}} \ge \gamma/2 \; , \\
\label{equation:strong-linear-separability-3}
& \begin{gathered}
\forall t \in \{1,2,\dots,T\} \quad \forall i \in \{1,2,\dots, K\} \setminus \{y_t\}, \qquad
\ip{x_t}{w_i} \le - \gamma/2 \; .
\end{gathered}
\end{align}
\end{enumerate}
\end{definition}

>>>>

\section{Notions of linear separability}
\label{section:notions-of-linear-separability}
Let $[n]=\{1,2,\ldots,n\}$.
We define two notions of linear separability for multiclass classification. The
first notion is the standard notion of linear separability used in the proof of
the mistake bound for the \textsc{Multiclass Perceptron} algorithm \citep[see e.g.][]{Crammer-Singer-2003}. The second
notion is stronger, i.e. more restrictive. %However, it is more suitable for the
%bandit setting, since it allows for a simple and efficient algorithm; see
%Section~\ref{section:algorithm-for-strongly-linearly-separable-data}.

\begin{definition}
[Linear separability]
\label{definition:linear-separability}
Let $(V,\ip{\cdot}{\cdot})$ be an inner product space, $K$ be a positive
integer, and $\gamma$ be a positive real number.
We say that labeled examples $(x_1, y_1),
(x_2, y_2), \dots, (x_T, y_T) \in V \times [K]$ are
\\ \vskip .01in
\emph{weakly linearly separable with a margin $\gamma$} if there exist vectors
$w_1, w_2, \dots, w_K \in V$ such that
\begin{align}
\label{equation:weak-linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1, \\
\label{equation:weak-linear-separability-2}
\ip{x_t}{w_{y_t}} & \ge \ip{x_t}{w_i} + \gamma \quad
\forall t \in [T]\ \forall i \in [K] \setminus \{y_t\}, 
\end{align}

and \emph{strongly linearly separable with a margin $\gamma$} if there exist vectors
$w_1, w_2, \dots, w_K \in V$ such that
\begin{align}
\label{equation:strong-linear-separability-1}
\sum_{i=1}^K \norm{w_i}^2 & \le 1, \\
\label{equation:strong-linear-separability-2}
\ip{x_t}{w_{y_t}} & \ge \gamma/2 \quad  \forall t \in [T], \\
\label{equation:strong-linear-separability-3}
\ip{x_t}{w_i} & \le - \gamma/2 \quad
\forall t \in [T]\ \forall i \in [K] \setminus \{y_t\}.
\end{align}
\end{definition}
